---
title: "_motifStack_: plot multiple motifs in one figure"
author: "Jianhong Ou^[Department of Cell Biology, Duke University School of Medicine; Regeneration Next Initiative, Durham, NC, USA], Michael H. Brodsky^[Department of Molecular, Cell and Cancer Biology, University of Massachusetts Medical School, Worcester, MA, USA], Lihua Julie Zhu^[Department of Molecular, Cell and Cancer Biology, University of Massachusetts Medical School, Worcester, MA, USA]"
date: "`r format(Sys.time(), '%B %d, %Y')`"
package: "`r packageVersion('motifStackWorkshop')`"
bibliography: "`r system.file('vignettes', 'ref.bib', package = 'motifStackWorkshop')`"
csl: "`r system.file('vignettes', 'nature.csl', package = 'motifStackWorkshop')`"
vignette: >
  %\VignetteIndexEntry{motifStack_workshop}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
output: 
  pdf_document:
    toc: yes
    toc_depth: 4
    fig_caption: true
  html_document:
    theme: simplex
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
---

```{r checkplatform, include=FALSE}
library(knitr)
suppressPackageStartupMessages({
  library(motifStack)
  library(MotifDb)
  library(ggplot2)
  library(RColorBrewer)
})
opts_chunk$set(eval=TRUE, fig.width=6, fig.height=2, warning = FALSE, message = FALSE)
extdata <- system.file("extdata", package = "motifStackWorkshop")
```

## Introduction

A sequence motif is a short sequence conservation in multiple amino acid or nucleic acid sequences
with biological significance such as DNA binding sites, where transcription factors (TFs) bind and catalytic sites in enzymes.
Many algorithms have been developed to discover and analyze motifs from DNA, 
RNA or protein sequences, which can be represented as position count matrix (PCM), position frequency matrix (PFM) 
or PWM (PWM) (reviewed in [@das2007survey;@sandve2007improved;@tompa2005assessing]). 
In turn, sequence logos, based on information theory or probability theory, 
are used as a graphical representation of sequence motifs by reflecting the frequencies 
of bases or amino acids at each conserved position[@schneider1990sequence;@schneider1996reading;@schneider2001strong;@crooks2004weblogo;@workman2005enologos].

High-throughput experimental approaches have been developed to generate motifs for large number of TFs in several organisms. 
The experimental method used to measure binding motifs including protein binding microarray (PBM)[@newburger2008uniprobe], bacterial one-hybrid (B1H)[@meng2005bacterial], systematic evolution of ligands by exponential enrichment (SELEX)[@ellington1990vitro;@tuerk1990systematic], High-Throughput (HT) SELEX[@hoinka2015large], selective microfluidics-based ligand enrichment followed by sequencing (SMiLE-seq)[@isakova2017smile], and so on. 
In downstream, different computational methods have been used to construct the motifs. 
MEME Suite[@bailey2009meme] is widely used tool for motifs construction for B1H and SELEX. 
Seed-and-Wobble (SW)[@badis2009diversity], Dream5[@weirauch2013evaluation] and BEEML[@zhao2011quantitative] are used for PBM data analysis.

There are many tools have been developed to represent the sequence motif as individual sequence logo, such as weblogo[@crooks2004weblogo], seqlogo[@bembom1seqlogo]. To facilitate comparison of motifs, more tools have been 
developed for motif alignment and clustering such as Tomtom[@gupta2007quantifying], MatAlign[@zhao2012conserved] and STAMP[@mahony2007stamp]. However, those tools are lacking flexibility in dispalying the similarities or differences among 
different groups of motifs.

The _motifStack_ package[@ou2018motifstack] is designed for graphic representation of multiple motifs for their similarities or differences.
It provides the flexibility for users to customize the graphic parameters such as the font family, colors, symbols, and highlight markers.

## Installation

_BiocManager_ is used to install stable version of _motifStack_.

```{r eval=FALSE}
library(BiocManager)
install("motifStack")
```

To install newest development version of _motifStack_, please try

```{r eval=FALSE}
install("jianhong/motifStack")
```

```{r}
packageVersion("motifStack")
```

Please note that start from version 1.28.0, _motifStack_ does not require ghostscript but cairo (>= 1.6).
If version number is smaller than 1.28.0, you will need ghostscript. The full path to the executable must be available 
for R session.

In this tutorial, version number greater than 1.29.4 is required.

## Inputs of _motifStack_

The inputs of _motifStack_ are one or more motifs represented as PCM or PFM. The PFM format motifs are available in
most of the motif databases such as TRANSFAC[@wingender2000transfac], JASPAR[@sandelin2004jaspar], CIS-BP[@weirauch2014determination] and HOCOMOCO[@kulakovskiy2017hocomoco]. More databases could be found at http://meme-suite.org/db/motifs.

`MotifDb` package is a collection of DNA-binding sequence motifs in PFM format. 9933 PFMs were collected from 11 sources: cisbp_1.02[@weirauch2014determination], FlyFactorSurvey[@zhu2010flyfactorsurvey], HOCOMOCO[@kulakovskiy2017hocomoco], HOMER[@heinz2010simple], hPDI[@xie2009hpdi], JASPAR[@sandelin2004jaspar],
Jolma2013[@jolma2013dna], ScerTF[@spivak2011scertf], stamlab[@neph2012circuitry], SwissRegulon[@pachkov2006swissregulon] and UniPROBE[@newburger2008uniprobe].

```{r}
library(MotifDb)
MotifDb[1]
MotifDb[[1]]
```

`plotMotifLogo` function be used to plot sequence logo for matrix in PFM format (Figure \ref{fig:plotMotifLogo}). 
As default, the nucleotides in a sequence logo will be drawn proportional to their information content (IC)[@workman2005enologos].
```{r fig.cap="\\label{fig:plotMotifLogo}Plot sequence logo for matrix in PFM format."}
library(motifStack)
plotMotifLogo(MotifDb[[1]])
```

The PFM can be formated as an `pfm` object and `pfm` object is the prefered format for _motifStack_ to plot (Figure \ref{fig:plotMotifPFM}).

```{r fig.cap="\\label{fig:plotMotifPFM}Plot sequence logo for `pfm` object."}
matrix.bin <- query(MotifDb, "bin_SANGER")
motif <- new("pfm", mat=matrix.bin[[1]], name=names(matrix.bin)[1])
motifStack(motif)
```

The `pfm` or `pcm` object can be created by PFM or PCM (Figure \ref{fig:plotMotifPCM}).

```{r fig.cap="\\label{fig:plotMotifPCM}Plot sequence logo for `pcm` object."}
path <- system.file("extdata", package = "motifStack")
pcm <- read.table(file.path(path, "bin_SOLEXA.pcm"))
pcm
pcm <- pcm[,3:ncol(pcm)]
rownames(pcm) <- c("A","C","G","T") # must have rownames
motif <- new("pcm", mat=as.matrix(pcm), name="bin_SOLEXA")
plot(motif)
```

`importMatrix` function can import the motifs into `pcm` or `pfm` object from files exported by TRANSFAC[@wingender2000transfac], JASPAR[@sandelin2004jaspar] and CIS-BP[@weirauch2014determination].

```{r,fig.height=8,fig.cap="\\label{fig:importMatrix}Plot sequence logo for data imported by `importMatrix`"}
motifs <- importMatrix(dir(path, "*.pcm", full.names = TRUE))
motifStack(motifs)
```


## Plot single sequence logo

_motifStack_ can be used to plot DNA/RNA sequence logo, affinity logo and amino acid sequence log. 
It provides the flexibility for users to customize the graphic parameters such as 
the font family, symbol colors and markers. 
Different from most of the sequence logo plotting tools, _motifStack_ does not pre-define the path 
for plotting. 
It depends on grImport/2[@murrell2009importing] package to import fonts into a path. 
This provides the power of _motifStack_ more flexibility to plot the logo.

## Plot a DNA/RNA/amino acid sequence logo

To plot a RNA sequence logo, change the rowname of the matrix from "T" to "U" (Figure \ref{fig:plotRNAmotif}).

```{r fig.cap="\\label{fig:plotRNAmotif}Plot RNA sequence logo."}
rna <- pcm
rownames(rna)[4] <- "U"
RNA_motif <- new("pcm", mat=as.matrix(rna), name="RNA_motif")
plot(RNA_motif)
```

If the PFM of amino acid sequence motif is provided, _motifStack_ can be used to draw amino acid sequence logo (Figure \ref{fig:plotAAmotif}).

```{r fig.cap="\\label{fig:plotAAmotif}Plot amino acid sequence logo."}
protein<-read.table(file.path(path,"cap.txt"))
protein<-t(protein[,1:20])
protein<-pcm2pfm(protein)
protein_motif<-new("pfm", mat=protein, name="CAP", 
                   color=colorset(alphabet="AA",colorScheme="chemistry"))
plot(protein_motif)
```

_motifStack_ can also be used to plot other logos (Figure \ref{fig:plotANYlogo}).

```{r fig.cap="\\label{fig:plotANYlogo}Plot logo of other types."}
m <- matrix(0, nrow = 10, ncol = 10, 
            dimnames = list(strsplit("motifStack", "")[[1]],
                            strsplit("motifStack", "")[[1]]))
for(i in seq.int(10)) m[i, i] <- 1
ms <- new("pfm", mat=m, name="motifStack")
plot(ms)
```

The sequence logo can be visulized with different fonts and colors (Figure \ref{fig:changeFont}).

```{r fig.cap="\\label{fig:changeFont}Plot sequence logo with different fonts and colors."}
motif$color <- colorset(colorScheme = 'blindnessSafe')
plot(motif, font="mono,Courier", fontface="plain")
```


### Plot sequence logo with highlight markers

`markers` slot of `pfm` or `pcm` object can be assigned by a list of `marker` objects. 
There are three type of markers, "rect", "line" and "text". 
If `markers` slot is assigned, markers can be plotted (Figure \ref{fig:withMarkers}).

```{r fig.cap="\\label{fig:withMarkers}Plot sequence logo with highlight markers."}
markerRect <- new("marker", type="rect", start=6, stop=7, 
                  gp=gpar(lty=2, fill=NA, col="orange", lwd=3))
markerRect2 <- new("marker", type="rect", start=2, 
                   gp=gpar(lty=3, fill=NA, col="green", lwd=2))
markerLine <- new("marker", type="line", start=2, stop=7,
                  gp=gpar(lwd=2, col="darkblue"))
markerText <- new("marker", type="text", start=c(1, 5), 
                  label=c("*", "core"), gp=gpar(cex=2, col="red"))
motif <- new("pcm", mat=as.matrix(pcm), name="bin_SOLEXA", 
             markers=c(markerRect, markerLine, markerText, markerRect2))
plot(motif)
```


## Plot motifs with ggplot2

Sequence logo could be plotted by `geom_motif` function which is compatible with _ggplot2_[@wickham2016ggplot2] (Figure \ref{fig:ggplot2}).

```{r,fig.height=8,fig.cap="\\label{fig:ggplot2}Plot sequence logo with ggplot2."}
len <- length(motifs)
motifs[[1]]$markers <- list(markerRect2, markerRect)
df <- data.frame(x=.5, y=(seq.int(len)-.5)/len, 
                 width=.75, height=1/(len+1))
df$motif <- motifs
library(ggplot2)
ggplot(df, aes(x=x, y=y, width=width, height=height, motif=motif)) +
  geom_motif(use.xy = TRUE) + theme_bw() + xlim(0, 1) + ylim(0, 1)
```


## Plot multiple sequence logos

_motifStack_ is designed to visulize the alignment of multiple motifs and facilitate the analysis
of binding site diversity/conservation within families of TFs and evolution of binding sites among
different species. There are many functions in _motifStack_ such as `plotMotifLogoStack`, `motifCloud`,
`plotMotifStackWithRadialPhylog`, `motifPiles`, and `motifCircos` for plotting motifs in different layouts,
`motifSignature` for generating motif signatures by merging similar motifs. 
To make it easy to use, we created one workflow function called `motifStack`, 
which can be used to draw a single DNA, RNA or amino acid motif as a sequence logo,
as well as align motifs, generate motif signatures and plot aligned motifs as a stack,
phylogeny tree, radial tree or cloud of sequence logos.

To align motifs, `DNAmotifAlignment` function implements Ungapped	Smith–Waterman	(local)	alignment 
based on column comparison scores calculated by Kullback-Leibler distance[@roepcke2005t]. 
`DNAmotifAlignment` requires the inputs be sorted by the distance of motifs. 
To get the sorted motifs for alignment, distances can be calculated using motifDistance in `motIV` 
(R implementation of STAMP[@mahony2007stamp]). 
Distances generated from MatAlign[@zhao2012conserved] are also supported with a helper function `ade4::newick2phylog`[@dray2007ade4]. 

### Plot sequence logo stack

As shown in above, `motifStack` function can be used to plot motifs as a stack of sequence logos easily.
By setting the layout to "tree" (Figure \ref{fig:treeLayout}) or "phylog" (Figure \ref{fig:phylogLayout}), 
`motifStack` will plot the stack of sequence logos with a linear tree.

```{r, fig.height=6, fig.cap="\\label{fig:treeLayout}Plot sequence logo stack with tree layout."}
motifStack(motifs, layout = "tree")
```
```{r, fig.height=6, fig.cap="\\label{fig:phylogLayout}Plot sequence logo stack with phylog layout."}
motifStack(motifs, layout="phylog", f.phylog=.15, f.logo=0.25)
```

The layout of 'phylog' has the capability to plot more motifs comparing to layout of 'tree'.
These plots allow small sets of motifs to be readily compared. 
However, moderate throughput analysis of TFs has generated 100s to 1000s of motifs, 
which make the visualization of their relationships more challenging. 
To reduce the total number of motifs depicted, clusters of highly similar motifs can be merged 
with the `motifSignature` function using a distance threshold. 
The distance threshold is user-determined; a low threshold will merge only extremely similar motifs
while a higher threshold can group motifs with lower similarity. 

```{r}
matrix.fly <- query(MotifDb, "FlyFactorSurvey")
motifs2 <- as.list(matrix.fly)
## format the name
names(motifs2) <- gsub("(_[\\.0-9]+)*_FBgn\\d+$", "", 
                       elementMetadata(matrix.fly)$providerName)
names(motifs2) <- gsub("[^a-zA-Z0-9]", "_", names(motifs2))

motifs2 <- motifs2[unique(names(motifs2))]
set.seed(1)
pfms <- sample(motifs2, 50)

## use MotIV to calculate the distances of motifs
jaspar.scores <- MotIV::readDBScores(file.path(find.package("MotIV"), 
                                               "extdata", 
                                               "jaspar2010_PCC_SWU.scores"))
d <- MotIV::motifDistances(lapply(pfms, pfm2pwm))
hc <- MotIV::motifHclust(d, method="average")
## convert the hclust to phylog object
phylog <- hclust2phylog(hc)
## reorder the pfms by the order of hclust
leaves <- names(phylog$leaves)
pfms <- pfms[leaves]
## create a list of pfm objects
pfms <- mapply(pfms, names(pfms), 
               FUN=function(.pfm, .name){
                 new("pfm",mat=.pfm, name=.name)})
## extract the motif signatures
motifSig <- motifSignature(pfms, phylog, groupDistance=0.01, min.freq=1)
```

Logos for these merged motifs can be plotted on the same dendrograms, 
using the relative size of the logos to reflect the number of TFs that contribute to the merged motif.
Compared with the original dendrogram, the number of TFs with similar motifs and 
the relationships between the merged motifs can be more easily compared.
The 50 motifs shown in the original image are now shown as only 28 motif clusters/signatures, 
with the members of each cluster clearly indicated. 
By setting the colors of dendrogram via parameter col.tree,
the colors of lable of leaves via parameter col.leaves,
and the colors of annotation bar via parameter col.anno, 
`motifPiles` function can mark each motif with multiple annotation colors (Figure \ref{fig:motifPiles}).

```{r,fig.height=8, fig.cap="\\label{fig:motifPiles}Plot sequence logo stack by motifPiles."}
## get the signatures from object of motifSignature
sig <- signatures(motifSig)
## get the group color for each signature
gpCol <- sigColor(motifSig)

library(RColorBrewer)
color <- brewer.pal(12, "Set3")
## plot the logo stack with pile style.
motifPiles(phylog=phylog, pfms=pfms, pfms2=sig, 
            col.tree=rep(color, each=5),
            col.leaves=rep(rev(color), each=5),
            col.pfms2=gpCol, 
            r.anno=c(0.02, 0.03, 0.04), 
            col.anno=list(sample(colors(), 50), 
                          sample(colors(), 50), 
                          sample(colors(), 50)),
            motifScale="logarithmic",
            plotIndex=TRUE,
            groupDistance=0.01)
```


### Plot sequence logo cloud

To emphasize the weight of motif signatures, motif signatures can be plotted as a circular (Figure \ref{fig:motifCloud}) or rectanglular 
word clouds with motif size representing the number of motifs that contributed to the signature. 
The larger the motif size, the larger the number of motifs within that motif signature.

```{r,fig.height=6, fig.cap="\\label{fig:motifCloud}Plot sequence logo cloud."}
## draw the motifs with a word-cloud style.
motifCloud(motifSig, scale=c(6, .5), layout="cloud")
```

### Plot sequence logo circos

The sequence logo clouds will loose the cluster information. 
Plotting motifs as sequence logo circos is a good solution to visualize motif signatures with dendrogram (Figure \ref{fig:plotMotifStackWithRadialPhylog}).

```{r,fig.height=6, fig.cap="\\label{fig:plotMotifStackWithRadialPhylog}Plot sequence logo circos."}
## plot the logo stack with radial style.
plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, 
                              circle=0.4, cleaves = 0.3, 
                              clabel.leaves = 0.5, 
                              col.bg=rep(color, each=5), col.bg.alpha=0.3, 
                              col.leaves=rep(rev(color), each=5),
                              col.inner.label.circle=gpCol, 
                              inner.label.circle.width=0.03,
                              angle=350, circle.motif=1.2, 
                              motifScale="logarithmic")
```

`motifCircos` function provide the possibility to plot the motif signatures with their original sequence logos
in the circos style (Figure \ref{fig:motifCircos}).

```{r,fig.height=6, fig.cap="\\label{fig:motifCircos}Plot sequence logo circos by motifCiros."}
## plot the logo stack with cirsoc style.
motifCircos(phylog=phylog, pfms=pfms, pfms2=sig, 
            col.tree.bg=rep(color, each=5), col.tree.bg.alpha=0.3, 
            col.leaves=rep(rev(color), each=5),
            col.inner.label.circle=gpCol, 
            inner.label.circle.width=0.03,
            col.outer.label.circle=gpCol, 
            outer.label.circle.width=0.03,
            r.rings=c(0.02, 0.03, 0.04), 
            col.rings=list(sample(colors(), 50), 
                           sample(colors(), 50), 
                           sample(colors(), 50)),
            angle=350, motifScale="logarithmic")
```

## Plot mouse data from different computational methods

Following scripts will show how _motifStack_ facilitated the discovery that changing the computational method used to build motifs from the same binding data can lead to artificial segregation of motifs for identical TFs within a motif alignment. 
MatAlign[@zhao2012conserved] was downloaded from http://stormo.wustl.edu/MatAlign/ and neighbor is a tool of [phylip](http://evolution.genetics.washington.edu/phylip.html). MatAlign2tree.pl is a perl script to run MatAlign and 
neighbor to get the neighbor-joint phylogenetic tree for given motifs. 
Compare to STAMP, MatAlign alignment method is superior for discriminating between the closely related motifs[@ou2018motifstack]. 

To confirm that the motif generating method was responsible for the apparent species dependent segregation of motifs, we clustered motifs for a smaller set of only mouse HDs that were determined using two different experimental methods to measure DNA binding specificity, SELEXseq and PBM, and different computational methods to generate the PBM motifs 
These results confirm that the motif generation method can impact clustering in several ways. First, the low IC motifs generated by one computational method (Dream5) tend to cluster with each other (motifs 31-56). Second, when many TFs have similar binding specificities by one method, the differences in motif generation may be the only source for differences between the motifs (motifs 1-30). Third, the SELEX motif set occasionally provides multiple motifs for a single TF; Lhx8 (highlighted in red) has a monomeric motif and two distinct dimeric motifs; only the monomeric motif clusters with the single motif generated by the PBM method.

Together, these results indicate that the experimental and computational methods for generating TF binding motifs can substantially alter the motifs generated for either a single TF or for clear orthologs. In some cases, these differences may reflect a genuine difference in the accurate modeling of the TF DNA binding specificities. However, it is clear that differences in the total IC and in the recovery of multiple motifs, including dimeric versus monomeric binding sites, will play a significant role in the clustering and comparison of motifs for the same TF. It is highly likely that they will also affect the ability to use these motifs to describe _in vivo_ function. Thus, different computational and experimental platforms must be carefully taken into account when searching for trends among large sets of TF motifs.

```{r, fig.width=12, fig.height=12, fig.cap="\\label{fig:MMdata}Mouse data from different methods."}
path <- system.file('libs', package = 'motifStackWorkshop')
pcmpath <- system.file('extdata', "pcmsDatasetMmPBM_SELEX", 
                       package = 'motifStackWorkshop')
MatAlign2tree <- system.file('extdata', 'MatAlign2tree.pl', 
                             package = 'motifStackWorkshop')

pcms <- readPCM(pcmpath)
pfms <- lapply(pcms, pcm2pfm)
outpath="output"
matalign_path <- file.path(path, "matalign")
neighbor_path <- file.path(path, "neighbor")
system(paste("perl", MatAlign2tree, "--in", getwd(), "--pcmpath", pcmpath, 
             "--out", outpath, "--matalign", matalign_path, 
             "--neighbor", neighbor_path, "--tree","UPGMA"))
newickstrUPGMA <- readLines(con=file.path(outpath, "NJ.matalign.distMX.nwk"))
phylog <- newick2phylog(newickstrUPGMA, FALSE)
leaves <- names(phylog$leaves)
motifs <- pfms[leaves]
groupDistance <- 2.5
trim <- 0.2
motifSig <- motifSignature(motifs, phylog, groupDistance=groupDistance, 
                           min.freq=1, trim=trim)
sig <- signatures(motifSig)
gpCol <- sigColor(motifSig)

##set methods color
colorSet <- c("Dm"="#40E0D0", "b1h"="#40E0D0", "sw"="#008080", "bml"="darkgreen", 
              "Mm"="brown", "MmDREAM"="blue", "Ms"="#F69156", 
              "Hs"="#D900D9")
methods <- factor(grepl("_SELEX", leaves) + grepl("_M\\d{4}", leaves))
levels(methods) <- c("Mm", "MmDREAM", "Ms")
##c(Mm     = "PBM/Uniprobe/Seed-And-Wobble", 
##  MmDREAM= "PBM/CIS-BP/DREAM5", 
##  Ms     = "SELEX/CIS-BP")
levels(methods) <- colorSet[levels(methods)]
methods <- as.character(methods)

leaveNames <- gsub("_(SELEX|PBM)", "", gsub("_(Jolma|Berger)", "", leaves))
leaveCol <- ifelse(grepl("Lhx8", leaveNames, ignore.case=TRUE), "#D55E00", "#000000")

## calculate average of top 8 position information content for each motif
icgp <- sapply(sapply(motifs, getIC), function(.ele)
  mean(sort(.ele, decreasing=TRUE)[1:min(8, length(.ele))]))
icgp.ranges <- range(icgp)
icgp <- cut(icgp, 10, labels=colorRampPalette(c("#009E73", "#000000", "#D55E00"))(10))
icgp.image <- as.raster(matrix(rev(levels(icgp)), ncol=1))
icgp <- as.character(icgp)

motifCircos(phylog=phylog, pfms=DNAmotifAlignment(motifs), 
            col.tree.bg=methods, col.tree.bg.alpha=.3, r.rings=c(.1, .1), 
            col.rings=list(icgp, rep("white", length(icgp))), 
            col.inner.label.circle=gpCol, inner.label.circle.width=0.05,
            labels.leaves=leaveNames, col.leaves=leaveCol, 
            cleaves=.1, r.tree=1, r.leaves=.5, clabel.leaves=.9, 
            motifScale="logarithmic", angle=358, groupDistance=groupDistance,
            plotIndex=TRUE,groupDistanceLineCol='#D55E00')

legend(-2.4, 2.4, legend=c("Mm/Uniprobe/PBM/Seed-And-Wobble", 
                           "Mm/CIS-BP/PBM/DREAM5", "Mm/CIS-BP/SELEX"), 
       fill=highlightCol(colorSet[c("Mm", "MmDREAM", "Ms")], .3), 
       border="white", lty=NULL, bty = "n")

text(2.05, 2.3, labels="Information Content")
rasterImage(icgp.image, 2, 1.8, 2.04, 2.2, interpolate = FALSE)
text(c(2.05, 2.05), c(1.8, 2.2), labels=format(icgp.ranges, digits=3), adj=0)

```


## References
